### 2017-02-04 Thesis/latex/paperB.tex
Bibliography.

There are a number of model equations which
allow the approximate description of the evolution of the free surface
without having to provide a complete solution of the fluid flow
below the surface.
%
In the present contribution, interest is focused on the derivation and evaluation of
a non-local water-wave model known as the Whitham equation.
We extend here the results of the work \cite{Moldabayev2015}
to the case where surface tension is taken into account.
The model equation under study is written as
%
\begin{equation}
\label{dimWhitham}
%	\eta_t+\frac{3}{2}\frac{c_0}{h_0}\eta\eta_x+K_{h_0} \! \ast \eta_{x}=0
	\eta_t + W \eta_x + \frac 32 \eta \eta_x = 0
	,
\end{equation}
%
where the convolution kernel of the operator
\(
	W \eta_x = w(-i \partial _x) \eta_x
	= \big( {\mathcal F}^{-1} w \big) * \eta_x
\)	
is given in terms of the Fourier transform by
%
\begin{equation}
\label{symbol}
%	\mathcal{F} \big( K_{h_0} \big) =
%	{\textstyle{ \sqrt{\frac{g\tanh (h_0\xi)}{\xi}}}}
	w(\xi) = \sqrt{ ( 1 + \varkappa \xi^2 ) \Sfrac{ \tanh(\xi) }{ \xi }}.
\end{equation}
%
Here it is assumed that the variables are suitably normalized so that
the gravitational acceleration, the undisturbed depth of the fluid
and the density are all unity.
The surface fluid tension is included here by means of the
capillarity parameter $\varkappa$ which is the inverse of the Bond number.
The convolution can be thought of as a Fourier multiplier operator,
and \eqref{symbol} represents the Fourier symbol of the operator. 
It is also convenient from the analytical point of view to
regard the Whitham operator $W = w(-i \partial _x)$
as an integral with respect to the spectral measure
of the self-adjoint operator $-i \partial _x$ in $L^2( \mathbb R )$.
Thus after linearisation \eqref{dimWhitham} can be considered
as a Schr\"odinger equation with the self-adjoint operator
$-i \partial _x w(-i \partial _x)$.
Indeed, introducing operator $D = -i \partial _x$ one may
rewrite \eqref{dimWhitham} as
%
\[
	i \eta_t = D w(D) \eta + \frac 34 D \eta^2
	.
\]
%
From this point of view, for example, one may deduce straight away that for any
real valued solution $\eta \in C^1( \mathbb R, L^2( \mathbb R ) )$
of this equation the $L^2$-norm does not depend on time.

The Whitham equation was proposed by Whitham \cite{Wh1} as an alternative to the
well known Korteweg-de Vries (KdV) equation
%
\begin{equation}
\label{KdV}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx} = 0
	.
\end{equation}
%
Provided $\varkappa < 1/3$ one may rescale $x$ and $t$ by
$\sqrt{1 - 3 \varkappa}$ and arrive at the equation
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac 16 \eta_{xxx} = 0
	.
\]
%
Thus it is apparent that small capillary effect do not add anything new
to the KdV model.
However, when $\varkappa$ is near $1/3$ one cannot expect
that this model be applicable.
To describe surface waves in such a situation, one may 
use instead the fifth-order-model equation
%
\begin{equation}
\label{Kawahara}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx}
	+ \frac{1}{360} ( 19 - 30 \varkappa - 45 \varkappa^2 ) \eta_{xxxxx}
	= 0
	.
\end{equation}
%
In our numerical experiments we use $\varkappa = 1/3$, so that the equation
reduces to what is known as the Kawahara equation \citep{Biswas2009, Kawahara1972}, which 
has the following form:
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac{1}{90} \eta_{xxxxx}
	= 0
	.
\]
%

@@

Both KdV and Kawahara equations are generally believed to approximate
very long waves quite well,
but one notorious problem with these equations is that they do not model
accurately the dynamics of shorter waves.
Recognizing this shortcoming of the KdV equation, Whitham proposed to use
the same nonlinearity as the KdV equation, but coupled with a linear term
which mimics the linear dispersion relation of the full water-wave problem.
Thus, at least in theory, 
the Whitham equation can be expected to yield a 
description of the dynamics of shorter waves which is closer to
the governing Euler equations.
The Whitham equation \eqref{dimWhitham} has been studied from
a number of vantage points during recent years.
In particular, the existence of traveling and solitary waves
has been studied  \citep{Minzoni2013, Ehrnstrom2012, Ehrnstrom2009, Ehrnstrom2013}.
Well posedness of a similar equation
was investigated in \cite{Lannes2013}.
Moreover, it has been shown
in \citep{Hur2015, Hur2015104,Sanford2014}
that periodic solutions of equation \eqref{dimWhitham}
feature modulational instability 
for short enough waves in a similar way as small-amplitude periodic wave
solutions of the water-wave problem. 
The performance of the Whitham equation in the description of surface water
waves has been investigated in \cite{BKN} in the steady case without surface tension.
However, it appears that no study of the performance of the Whitham equation
in the presence of capillarity has been done.

@@

The total energy of the system is given by the sum of the kinetic energy,
the potential energy and the surface tension energy,
and normalized in such a way that the total energy
is zero when no wave motion is present at the surface.
Accordingly the Hamiltonian function for this problem is
%
\[
	H = \int _{\mathbb R} \int_0^\eta z \, dz dx +
	\int _{\mathbb R} \int_{-1}^\eta \Sfrac{1}{2} |\nabla \phi|^2 \, dz dx
	+ \varkappa \int _{\mathbb R} \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} } dx	
	.
\]
%
Defining the trace of the potential at the free surface as
$\Phi(x,t) = \phi(x,\eta(x,t),t)$,
one may integrate in $z$ in the first integral and use the divergence theorem 
on the second integral in order to arrive at the formulation
%
\begin{equation}
\label{Hamiltonian-eta-phi}
	H  = \int_\R \Big[ \Sfrac{1}{2} \eta^2 + \Sfrac{1}{2} \Phi G(\eta) \Phi
	+ \varkappa \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} }	
	\Big] \, dx.
\end{equation}
%
This is the Hamiltonian of the water wave problem with surface tension
as for instance found in \cite{Alazard2011},
and written in terms of the Dirichlet-Neumann operator $G(\eta)$.
As shown in \cite{Nicholls}, the Dirichlet-Neumann operator is analytic
in a certain sense, and can be expanded as a power series as
%
\[
	G(\eta)\Phi = \sum_{j=0}^\infty G_j(\eta) \Phi
	.
\]
%
In order to proceed, we need to understand the first few terms in this series.
As shown in \cite{CS} and \cite{CG}, the first two terms in this series
can be written with the help of the operator $D = - i \partial_x$ as 
%
\[
	G_0(\eta)= D\tanh(D)
	, \qquad 
	G_1(\eta)= D\eta D - D\tanh(D) \eta D\tanh(D)
	. 
\]
%
%
Note that it can be shown that the terms $G_j(\eta)$ for $j\ge 2$ are of quadratic or
higher-order in $\eta$, and will therefore not be needed in the following
analysis.

@@

Now the Hamiltonian (\ref{Hamiltonian_eta_u}) may be simplified as follows

@@

Making reference to \citep{Craig1994}, Craig2005} note that

@@

\subsection{Linear approximation.}

@@

\subsection{The shallow-water scaling regime.}

@@

\subsection{The Boussinesq scaling regime.}

>>>>

There are a number of model equations which
allow the approximate description of the evolution of the free surface
without having to provide a complete solution of the fluid flow
below the surface.
%
In the present contribution, interest is focused on the derivation and evaluation of
a non-local water-wave model known as the Whitham equation.
We extend here the results of the work \cite{Moldabayev2015}
to the case where surface tension is taken into account.
The model equation under study is written as
%
\begin{equation}
\label{dimWhitham}
%	\eta_t+\frac{3}{2}\frac{c_0}{h_0}\eta\eta_x+K_{h_0} \! \ast \eta_{x}=0
	\eta_t + W \eta_x + \frac 32 \eta \eta_x = 0
	,
\end{equation}
%
where the convolution kernel of the operator
\[
	W \eta_x = w(-i \partial _x) \eta_x
	= \big( {\mathcal F}^{-1} w \big) * \eta_x
\]	
is given in terms of the Fourier transform by
%	\mathcal{F} \big( K_{h_0} \big) =
%	{\textstyle{ \sqrt{\frac{g\tanh (h_0\xi)}{\xi}}}}
%
\begin{equation}
\label{symbol}
	w(\xi) = \sqrt{ ( 1 + \varkappa \xi^2 ) \Sfrac{ \tanh(\xi) }{ \xi }}.
\end{equation}
%
Here it is assumed that the variables are suitably normalized so that
the gravitational acceleration, the undisturbed depth of the fluid
and the density are all unity.
The surface fluid tension is included here by means of the
capillarity parameter $\varkappa$ which is the inverse of the Bond number.
The convolution can be thought of as a Fourier multiplier operator,
and \eqref{symbol} represents the Fourier symbol of the operator. 
It is also convenient from the analytical point of view to
regard the Whitham operator $W = w(-i \partial _x)$
as an integral with respect to the spectral measure
of the self-adjoint operator $-i \partial _x$ in $L^2( \mathbb R )$.
Thus after linearisation \eqref{dimWhitham} can be considered
as a Schr\"odinger equation with the self-adjoint operator
$-i \partial _x w(-i \partial _x)$.
Indeed, introducing operator $D = -i \partial _x$ one may
rewrite \eqref{dimWhitham} as
%
\[
	i \eta_t = D w(D) \eta + \frac 34 D \eta^2
	.
\]
%
From this point of view, for example, one may deduce straight away that for any
real valued solution $\eta \in C^1( \mathbb R, L^2( \mathbb R ) )$
of this equation the $L^2$-norm does not depend on time.

The Whitham equation was proposed by Whitham \cite{Whitham1967} as an alternative to the
well known Korteweg-de Vries (KdV) equation
%
\begin{equation}
\label{KdV}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx} = 0
	.
\end{equation}
%
Provided $\varkappa < 1/3$ one may rescale $x$ and $t$ by
$\sqrt{1 - 3 \varkappa}$ and arrive at the equation
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac 16 \eta_{xxx} = 0
	.
\]
%
Thus it is apparent that small capillary effect do not add anything new
to the KdV model.
However, when $\varkappa$ is near $1/3$ one cannot expect
that this model be applicable.
To describe surface waves in such a situation, one may 
use instead the fifth-order-model equation
%
\begin{equation}
\label{Kawahara}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx}
	+ \frac{1}{360} ( 19 - 30 \varkappa - 45 \varkappa^2 ) \eta_{xxxxx}
	= 0
	.
\end{equation}
%
In our numerical experiments we use $\varkappa = 1/3$, so that the equation
reduces to what is known as the Kawahara equation \citep{Biswas2009, Kawahara1972}, which 
has the following form:
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac{1}{90} \eta_{xxxxx}
	= 0
	.
\]
%

@@

Both KdV and Kawahara equations are generally believed to approximate
very long waves quite well,
but one notorious problem with these equations is that they do not model
accurately the dynamics of shorter waves.
Recognizing this shortcoming of the KdV equation, Whitham proposed to use
the same nonlinearity as the KdV equation, but coupled with a linear term
which mimics the linear dispersion relation of the full water-wave problem.
Thus, at least in theory, 
the Whitham equation can be expected to yield a 
description of the dynamics of shorter waves which is closer to
the governing Euler equations.
The Whitham equation \eqref{dimWhitham} has been studied from
a number of vantage points during recent years.
In particular, the existence of traveling and solitary waves
has been studied  \citep{Minzoni2013, Ehrnstrom2012, Ehrnstrom2009, Ehrnstrom2013}.
Well posedness of a similar equation
was investigated in \cite{Lannes2013}.
Moreover, it has been shown
in \citep{Hur2015, Hur2015104,Sanford2014}
that periodic solutions of equation \eqref{dimWhitham}
feature modulational instability 
for short enough waves in a similar way as small-amplitude periodic wave
solutions of the water-wave problem. 
The performance of the Whitham equation in the description of surface water
waves has been investigated in \cite{Borluk2016} in the steady case without surface tension.
However, it appears that no study of the performance of the Whitham equation
in the presence of capillarity has been done.

@@

The total energy of the system is given by the sum of the kinetic energy,
the potential energy and the surface tension energy,
and normalized in such a way that the total energy
is zero when no wave motion is present at the surface.
Accordingly the Hamiltonian function for this problem is
%
\[
	H = \int _{\mathbb R} \int_0^\eta z \, dz dx +
	\int _{\mathbb R} \int_{-1}^\eta \Sfrac{1}{2} |\nabla \phi|^2 \, dz dx
	+ \varkappa \int _{\mathbb R} \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} } dx	
	.
\]
%
Defining the trace of the potential at the free surface as
$\Phi(x,t) = \phi(x,\eta(x,t),t)$,
one may integrate in $z$ in the first integral and use the divergence theorem 
on the second integral in order to arrive at the formulation
%
\begin{equation}
\label{Hamiltonian-eta-phi}
	H  = \int_\R \Big[ \Sfrac{1}{2} \eta^2 + \Sfrac{1}{2} \Phi G(\eta) \Phi
	+ \varkappa \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} }	
	\Big] \, dx.
\end{equation}
%
This is the Hamiltonian of the water wave problem with surface tension
as for instance found in \cite{Alazard2011},
and written in terms of the Dirichlet-Neumann operator $G(\eta)$.
As shown in \cite{Nicholls2001}, the Dirichlet-Neumann operator is analytic
in a certain sense, and can be expanded as a power series as
%
\[
	G(\eta)\Phi = \sum_{j=0}^\infty G_j(\eta) \Phi
	.
\]
%
In order to proceed, we need to understand the first few terms in this series.
As shown in \cite{Craig1993} and \cite{Craig1994}, the first two terms in this series
can be written with the help of the operator $D = - i \partial_x$ as 
%
\[
	G_0(\eta)= D\tanh(D)
	, \qquad 
	G_1(\eta)= D\eta D - D\tanh(D) \eta D\tanh(D)
	. 
\]
%
%
Note that it can be shown that the terms $G_j(\eta)$ for $j\ge 2$ are of quadratic or
higher-order in $\eta$, and will therefore not be needed in the following
analysis.

@@

Now the Hamiltonian \eqref{Hamiltonian_eta_u} may be simplified as follows

@@

Making reference to \citep{Craig1994, Craig2005} note that

@@

\subsubsection{Linear approximation.}

@@

\subsubsection{The shallow-water scaling regime.}

@@

\subsubsection{The Boussinesq scaling regime.}
